---
title: "Language Sentiment Predicts Changes in Depression (Manuscript Results Analyses)"
author: "Jihyun Hur"
output: html_document
---

```{r setup, include = FALSE}
library(dplyr)
library(stringr)
library(ggcorrplot)
library(ggpubr)
library(ggeffects)
library(jtools)
library(broom)
library(purrr)
library(car)
library(robustbase)
library(knitr)
library(kableExtra)
library(broom)
library(sjPlot)
library(psych)
library(gtsummary)
library(robustlmm)
library(tidyr)
library(lmerTest)
library(rcartocolor)  
library(rstatix)     
library(R.matlab)
library(ggtext)

# set working directory
cur_dir <- '/Users/jihyunhur/Yale/2_Github/project_language_sentiment_depression/scripts'
dat_dir <- '/Users/jihyunhur/Yale/2_Github/project_language_sentiment_depression/data'
dir_save <- '/Users/jihyunhur/Yale/2_Github/project_language_sentiment_depression/scripts/figures'

study1_data <- read.csv(file.path(dat_dir,'study1_data.csv'))
study2_data <- read.csv(file.path(dat_dir,'study2_data.csv'))
```

## Results
* The scripts for all analyses in the results section are included below in the order reported in the paper. 

Positive and negative human sentiment Spearman correlation
```{r ms 1, warning = FALSE}
## Study 1
cor.test(study1_data$avg_human_tone_pos_mean, study1_data$avg_human_tone_neg_mean, method="spearman")

## Study 2
cor.test(study2_data$avg_human_tone_pos_mean, study2_data$avg_human_tone_neg_mean, method="spearman")
```

Human sentiment score ~ PHQ-9 at baseline
```{r ms 2, warning = FALSE}
## Study 1
cor.test(study1_data$phq, study1_data$avg_human_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$phq, study2_data$avg_human_tone_mean_diff, method="spearman")
```

Human sentiment score ~ PHQ-9 at follow-up
```{r ms 3, echo = TRUE, warning = FALSE}
## Study 1
cor.test(study1_data$phq_2, study1_data$avg_human_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$phq_2, study2_data$avg_human_tone_mean_diff, method="spearman")
```

PHQ-9 at baseline ~ PHQ-9 at follow-up
```{r ms 4, echo = TRUE, warning = FALSE}
## Study 1
cor.test(study1_data$phq, study1_data$phq_2, method="spearman")

## Study 2
cor.test(study2_data$phq, study2_data$phq_2, method="spearman")
```

Robust regression testing human sentiment score 
```{r ms 5-1, echo = TRUE, warning = FALSE}
# define variable full name for regression result table (tab_model below)
pi <- c(`(Intercept)` = "Intercept",
        phq = "Initial PHQ-9 score",
        avg_human_tone_mean_diff = "Human sentiment score", 
        avg_human_tone_pos_mean = "Human positivity score",  
        avg_human_tone_neg_mean = "Human negativity score",
        gpt_tone_mean_diff = "ChatGPT sentiment score", 
        gpt_tone_pos_mean = "ChatGPT positivity score",
        gpt_tone_neg_mean = "ChatGPT negativity score",
        liwc_tone_mean_diff = "LIWC sentiment score", 
        liwc_tone_pos_mean = "LIWC positivity score", 
        liwc_tone_neg_mean = "LIWC negativity score",
        base_hap = "Baseline mood")

# define variables to inclde in regression result table (tab_model below)
terms_use <- c("phq", "avg_human_tone_mean_diff", "avg_human_tone_pos_mean", "avg_human_tone_neg_mean", 
               "gpt_tone_mean_diff", "gpt_tone_pos_mean", "gpt_tone_neg_mean",
               "liwc_tone_mean_diff", "liwc_tone_pos_mean", "liwc_tone_neg_mean",
               "base_hap")

## Study 1
ms5_study1_mod <- lmrob(phq_2   ~   phq + age + gender + edu + avg_human_tone_mean_diff, data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms5_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms5_study1_mod2<-lmrob(phq_2   ~  phq  + age + gender + edu, data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))

s1_mod_f <- (summary(ms5_study1_mod)$r.squared - summary(ms5_study1_mod2)$r.squared)/(1-summary(ms5_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms5_study2_mod <- lmrob(phq_2   ~  phq + age + gender + edu + avg_human_tone_mean_diff, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms5_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms5_study2_mod2<-lmrob(phq_2   ~  phq  + age + gender + edu, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))

s2_mod_f <- (summary(ms5_study2_mod)$r.squared - summary(ms5_study2_mod2)$r.squared)/(1-summary(ms5_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

Robust regression human sentiment plots (Fig. 1A)
```{r ms 5-2, warning = FALSE}
reg_mod <- plot_summs(ms5_study1_mod, ms5_study2_mod, scale=FALSE, inner_ci_level = .95,  plot.distributions = FALSE,
                 rescale.distributions = TRUE,  
                 coefs = c(
                   "Initial \nPHQ-9 score" = 'phq',
                   "Human \nsentiment score" = "avg_human_tone_mean_diff"),
                 model.names=c('Study 1', 'Study 2')) + scale_x_continuous(limits = c(-0.51, 1.01), 
                                                                           breaks=c(-0.50, 0.00, 0.50, 1.00))
fig_1a <- ggpar(reg_mod, xlab = 'Beta Coefficient',
            font.y = 0,
            ggtheme = theme_pubr()) + theme(legend.title = element_blank(), legend.position = c(0.3,0.95),
                                            legend.direction = "horizontal")
# print Fig. 1A
fig_1a
```

Robust linear regression model (rlm) prediction (Study 1)
```{r ms 5-3, echo = FALSE, eval = TRUE, warning = FALSE}
## Study 1 prediction using rlm
pred_values1 <- NULL

# Note: If only one participant in the category, the prediction model for that participant doesn't work
# exclude groups with only one participant
study1_data_pred <- study1_data %>% filter(!is.nan(age) & edu !="No high school diploma" & gender != "Others") 

for (i in 1:nrow(study1_data_pred)) {
  # sentiment model
  model1<-lmrob(phq_diff  ~ phq + avg_human_tone_mean_diff + age + gender + edu, data=study1_data_pred,  control = lmrob.control(setting = "KS2014", k.max=2000))
  pred_values1[i]  <- predict(model1, newdata=study1_data_pred[i,], type = "response")
}

study1_data_pred$pred_values         <- pred_values1

## Study 2 prediction using rlm
pred_values2 <- NULL

# Note: If only one participant in the category, the prediction model for that participant doesn't work
# exclude groups with only one participant
study2_data_pred <- study2_data %>% filter(!is.nan(age) & edu != "Prefer not to say" & gender != "Prefer not to say")

for (i in 1:nrow(study2_data_pred)) {
  # sentiment model
  model1<-lmrob(phq_diff   ~ phq + avg_human_tone_mean_diff + age + gender + edu, data=study2_data_pred,  control = lmrob.control(setting = "KS2014", k.max=2000))
  pred_values2[i] <- predict(model1, newdata=study2_data_pred[i,], type = "response")
}

study2_data_pred$pred_values         <- pred_values2

# Fig. 1B
merged_pred_dat <- full_join(study1_data_pred, study2_data_pred)

wilcox_tests = merged_pred_dat %>%
  filter(!is.na(phq_2) & !is.na(pred_values)) %>%
  group_by(study_num,
           Group = cut(pred_values, breaks = c(-Inf, -0.50, 0.50, +Inf),
           labels = c("Predicted \n PHQ-9 decrease", "Predicted \nPHQ-9 no change", "Predicted \nPHQ-9 increase"))) %>%
            summarise(P = wilcox.test(phq_diff, mu = 0)$p.value,
                    Sig = case_when(P < 0.005 ~ "***",
                                    P >= 0.05 ~ "ns",
                                    0.01 <= P & P < 0.05 ~ "*",
                                    0.005 <= P & P < 0.01 ~ "**"))
# print wilcox test
wilcox_tests

fig_1b <- merged_pred_dat %>%
  filter(!is.nan(phq_2) & !is.na(pred_values)) %>%
  group_by(study_num,
           Group = cut(pred_values, breaks = c(-Inf, -0.50, 0.50, +Inf),
           labels = c("Predicted \nPHQ-9 decrease", "Predicted \nPHQ-9 no change", "Predicted \nPHQ-9 increase"))) %>%
  summarise(mean_fit_value = mean(pred_values), sem_fit_value = sd(pred_values) / sqrt(n()),
            mean_real_value = mean(phq_diff), sem_real_value = sd(phq_diff) / sqrt(n())) %>%
  ggplot(aes(y = Group, x = mean_real_value)) +
  geom_vline(xintercept=0, alpha=0.5, linetype="dotted") +
  geom_pointrange(aes(y = Group, xmin = mean_real_value - sem_real_value, xmax = mean_real_value + sem_real_value,
  color = factor(study_num), shape = factor(study_num)), position = ggplot2::position_dodge(width = -0.6)) +
                    scale_shape_manual(values=c(1, 0), labels = c("Study 1", "Study 2")) +
                    scale_color_manual(values=c("#49b7fc", "#ff7b00"), labels = c("Study 1", "Study 2"))+
  labs(x = "Actual PHQ-9 Change", y="") + theme_pubr() + theme(legend.position = c(0.3,0.95), legend.direction = "horizontal", legend.title=element_blank())

# print Fig. 1B
fig_1b

# save Fig. 1
fig_1 <- ggarrange(fig_1a, fig_1b, ncol = 2, labels = c("A", "B"))
ggsave(filename = file.path(dir_save, "fig_1.pdf"), plot=fig_1, width=7, height=3.5, units="in", dpi=700)

```

LOOCV prediction (Study 1)
```{r ms 6-1, warning = FALSE}
## Study 1 LOOCV prediction
loocv_pred_values1 <- NULL

# Note: If only one participant in the category, the loocv model for that participant doesn't work
# exclude groups with only one participant
study1_data_loocv <- study1_data %>% filter(!is.nan(age) & edu !="No high school diploma" & gender != "Others") 

for (i in 1:nrow(study1_data_loocv)) {
  loo_data <- study1_data_loocv[-i,]

  # sentiment model
  model1<-lmrob(phq_diff  ~ phq + avg_human_tone_mean_diff + age + gender + edu, data=loo_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
  loocv_pred_values1[i]  <- predict(model1, newdata=study1_data_loocv[i,], type = "response")
}

study1_data_loocv$loocv_pred_values         <- loocv_pred_values1
```

LOOCV prediction (Study 2)
```{r ms 6-2, warning = FALSE}
## Study 2 LOOCV prediction
loocv_pred_values2 <- NULL

# Note: If only one participant in the category, the loocv model for that participant doesn't work
# exclude groups with only one participant
study2_data_loocv <- study2_data %>% filter(!is.nan(age) & edu != "Prefer not to say" & gender != "Prefer not to say")

for (i in 1:nrow(study2_data_loocv)) {
  loo_data <- study2_data_loocv[-i,]
  # sentiment model
  model1<-lmrob(phq_diff   ~ phq + avg_human_tone_mean_diff + age + gender + edu, data=loo_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
  loocv_pred_values2[i] <- predict(model1, newdata=study2_data_loocv[i,], type = "response")
}

study2_data_loocv$loocv_pred_values         <- loocv_pred_values2
```

LOOCV results 
```{r ms 6-3, warning = FALSE}
## Study 1
s1_mse <- mean((study1_data_loocv$phq_diff-study1_data_loocv$loocv_pred_values)^2, na.rm=TRUE)
sprintf("Study 1 LOOCV RMSE = %.2f", sqrt(s1_mse))

s1_mae <- mean(abs(study1_data_loocv$phq_diff-study1_data_loocv$loocv_pred_values), na.rm=TRUE)
sprintf("Study 1 LOOCV MAE = %.2f", s1_mae)

s1_R <- cor.test(study1_data_loocv$phq_diff, study1_data_loocv$loocv_pred_values, method="pearson")
sprintf("Study 1 Pred ~ Actual R = %.2f (P = %.3f)", s1_R$estimate[[1]], s1_R$p.value)

## Study 2
s2_mse <- mean((study2_data_loocv$phq_diff-study2_data_loocv$loocv_pred_values)^2, na.rm=TRUE)
sprintf("Study 2 LOOCV RMSE = %.2f", sqrt(s2_mse))

s2_mae <- mean(abs(study2_data_loocv$phq_diff-study2_data_loocv$loocv_pred_values), na.rm=TRUE)
sprintf("Study 2 LOOCV MAE = %.2f", s2_mae)

s2_R <- cor.test(study2_data_loocv$phq_diff, study2_data_loocv$loocv_pred_values, method="pearson")
sprintf("Study 2 Pred ~ Actual R = %.2f (P = %.3f)", s2_R$estimate[[1]], s2_R$p.value)
```

Model comparison using Wald tests
```{r ms 6-4, include=TRUE, warning = FALSE}
## Study 1
anova(ms5_study1_mod, ms5_study1_mod2, test='Wald')

## Study 2
anova(ms5_study2_mod, ms5_study2_mod2, test='Wald')
```

Robust mixed effect models to minimal and mild-to-moderate depression groups
```{r ms 6-5, include = TRUE, warning = FALSE}
# combine study1 and study2 data
dat <- rbind(study1_data, study2_data)

# mixed linear effect model for participants with minimal depression symptoms (PHQ-9 < 5) 
low_phq_lmer <- rlmer(phq_2   ~  phq + age + gender + edu + avg_human_tone_mean_diff + (1 |study_num), data = dat %>% filter(phq<5))

tab_model(low_phq_lmer, terms = terms_use, pred.labels = pi, dv.labels = c(""), 
          show.ci = FALSE,  show.se = TRUE, show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat", 
          string.se = "SE", string.p = "P")

# mixed linear effect model for participants with minimal depression symptoms (PHQ-9 >= 5)
high_phq_lmer <- rlmer(phq_2   ~  phq + age + gender + edu +  avg_human_tone_mean_diff + (1 |study_num), data = dat %>% filter(phq>=5))

tab_model(high_phq_lmer, terms = terms_use, pred.labels = pi, dv.labels = c(""), 
          show.ci = FALSE, show.se = TRUE, show.stat = TRUE, string.est = "Beta estimates", 
          string.stat = "T-stat", string.se = "SE", string.p = "P")
```

Linguistic distance ~ PHQ
```{r ms 7, warning = FALSE}
## Study 1
cor.test(study1_data$liwc_ling_dist, study1_data$phq, method="spearman")

# Study 2
cor.test(study2_data$liwc_ling_dist, study2_data$phq, method="spearman")
```

Correlation between human and ChatGPT (GPT-3.5) sentiment
```{r ms 8, warning = FALSE}
## Study 1
cor.test(study1_data$avg_human_tone_mean_diff, study1_data$gpt_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$avg_human_tone_mean_diff, study2_data$gpt_tone_mean_diff, method="spearman")
```

Correlation between PHQ-9 and ChatGPT (GPT-3.5) sentiment
```{r ms 9, warning = FALSE}
## Study 1
cor.test(study1_data$phq, study1_data$gpt_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$phq, study2_data$gpt_tone_mean_diff, method="spearman")
```

Robust regression testing ChatGPT sentiment score 
```{r ms 10-1, warning = FALSE}
## Study 1
ms10_study1_mod <- lmrob(phq_2   ~  phq + age + gender + edu +  gpt_tone_mean_diff, data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms10_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
            pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms10_study1_mod2<-lmrob(phq_2   ~  phq  + age + gender + edu, data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms10_study1_mod)$r.squared - summary(ms10_study1_mod2)$r.squared)/(1-summary(ms10_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms10_study2_mod <- lmrob(phq_2   ~  phq + age + gender + edu +  gpt_tone_mean_diff, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms10_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms10_study2_mod2<-lmrob(phq_2   ~  phq  + age + gender + edu, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms10_study2_mod)$r.squared - summary(ms10_study2_mod2)$r.squared)/(1-summary(ms10_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

Robust regression ChatGPT sentiment score plot (Fig. 2A)
```{r ms 10-2, warning = FALSE}
gpt_reg_mod <- plot_summs(ms10_study1_mod, ms10_study2_mod, scale=FALSE, inner_ci_level = .95,  plot.distributions = FALSE,
                 rescale.distributions = TRUE,   
                 coefs = c(
                   "Initial \nPHQ-9 score" = 'phq',
                   "ChatGPT \nsentiment score" = "gpt_tone_mean_diff"),
                 model.names=c('Study 1', 'Study 2')) + scale_x_continuous(limits = c(-0.51, 1.01), 
                                                                           breaks=c(-0.50, 0.00, 0.50, 1.00))
fig_2a <- ggpar(gpt_reg_mod, xlab = 'Beta Coefficient',
            font.y = 0,
            ggtheme = theme_pubr()) + theme(legend.title = element_blank(), legend.position = c(0.3,0.95),
                                            legend.direction = "horizontal")
# print Fig. 2A
fig_2a
```

LIWC ~ Human sentiment score
```{r ms 11, warning = FALSE}
## Study 1
cor.test(study1_data$liwc_tone_mean_diff, study1_data$avg_human_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$liwc_tone_mean_diff, study2_data$avg_human_tone_mean_diff, method="spearman")
```

LIWC ~ GPT sentiment score
```{r ms 12, warning = FALSE}
## Study 1
cor.test(study1_data$liwc_tone_mean_diff, study1_data$gpt_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$liwc_tone_mean_diff, study2_data$gpt_tone_mean_diff, method="spearman")
```

LIWC ~ PHQ-9 at baseline
```{r ms 13, warning = FALSE}
## Study 1
cor.test(study1_data$liwc_tone_mean_diff, study1_data$phq, method="spearman")

## Study 2
cor.test(study2_data$liwc_tone_mean_diff, study2_data$phq, method="spearman")
```

Robust regression testing LIWC sentiment score 
```{r ms 14-1, warning = FALSE}
## Study 1
ms14_study1_mod <- lmrob(phq_2   ~  phq + age + gender + edu + liwc_tone_mean_diff, data=study1_data,
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms14_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms14_study1_mod2<-lmrob(phq_2   ~  phq  + age + gender + edu, data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms14_study1_mod)$r.squared - summary(ms14_study1_mod2)$r.squared)/(1-summary(ms14_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms14_study2_mod <- lmrob(phq_2   ~  phq + age + gender + edu + liwc_tone_mean_diff, data=study2_data, 
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms14_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms14_study2_mod2<-lmrob(phq_2   ~  phq  + age + gender + edu, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms14_study2_mod)$r.squared - summary(ms14_study2_mod2)$r.squared)/(1-summary(ms14_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

Robust regression LIWC sentiment score plot (Fig. 2B)
```{r ms 14-2, warning = FALSE}
liwc_reg_mod <- plot_summs(ms14_study1_mod, ms14_study2_mod, scale=FALSE, inner_ci_level = .95,  plot.distributions = FALSE,
                 rescale.distributions = TRUE,   
                 coefs = c(
                   "Initial \nPHQ-9 score" = 'phq',
                   "LIWC \nsentiment score" = "liwc_tone_mean_diff"),
                 model.names=c('Study 1', 'Study 2')) + scale_x_continuous(limits = c(-0.51, 1.01), 
                                                                           breaks=c(-0.50, 0.00, 0.50, 1.00))
fig_2b <- ggpar(liwc_reg_mod, xlab = 'Beta Coefficient',
            font.y = 0,
            ggtheme = theme_pubr()) + theme(legend.title = element_blank(), legend.position = c(0.3,0.95),
                                            legend.direction = "horizontal")
# print Fig. 2B
fig_2b

# save Fig. 2B
fig_2 <- ggarrange(fig_2a, fig_2b, ncol = 2, nrow = 1, labels = c("A", "B")) 
ggsave(filename = file.path(dir_save, "fig_2.pdf"), plot=fig_2, width=7, height=3.5, units="in", dpi=700)
```

Mean r2 of happiness modeling for each study & baseline mood parameter ~ human sentiment score
```{r ms 15, warning = FALSE}
## Study 1 mean r2
s1_mat_file <- "study1_riskytask_data.mat"
s1_task_dat <- readMat(file.path(dat_dir, s1_mat_file))

all_s1_r2 <- NULL

for (i in 1:length(s1_task_dat$datatosave)) {
  all_s1_r2[i] <- s1_task_dat$datatosave[[i]][[13]]
}
all_s1_r2[!is.finite(all_s1_r2)] <- NA
# include participants who completed the study
mean(all_s1_r2[!is.na(study1_data$phq_2)], na.rm=TRUE)

## Study 2 mean r2
s2_mat_file <- "study2_riskytask_data.mat"
s2_task_dat <- readMat(file.path(dat_dir, s2_mat_file))

all_s2_r2 <- NULL

for (i in 1:length(s2_task_dat$datatosave)) {
  all_s2_r2[i] <- s2_task_dat$datatosave[[i]][[13]]
}

all_s2_r2[!is.finite(all_s2_r2)] <- NA
# include participants who completed the study
mean(all_s2_r2[!is.na(study2_data$phq_2)], na.rm=TRUE)

## Study 1 cor: mood par ~ sentiment
cor.test(study1_data$base_hap, study1_data$avg_human_tone_mean_diff, method="spearman")

## Study 2 cor: mood par ~ sentiment
cor.test(study2_data$base_hap, study2_data$avg_human_tone_mean_diff, method="spearman")
```

RPE ~ human sentiment score
```{r ms 16, warning = FALSE}
## Study 1
cor.test(study1_data$rpe, study1_data$avg_human_tone_mean_diff, method="spearman")

## Study 2
cor.test(study2_data$rpe, study2_data$avg_human_tone_mean_diff, method="spearman")
```

Risky task happiness plot (Fig. 3)
```{r ms 17, warning = FALSE}
# access the structure of Study 2 happiness data
main_struct <- s2_task_dat$datatosave
nrow <- dim(s2_task_dat$datatosave)[2]
df_list <- list()

# iterate through the rows
for (i in 1:nrow) {
  # Access the 1x1 struct for each row
  sub_struct <- main_struct[[i]]

  # Access and print the variables you need
  variable1 <- sub_struct[[1]] # rawhappy
  variable2 <- sub_struct[[12]] # Replace "variable2_name" with the actual variable name

  tmp_df <- data.frame(subjID = i, rawhappy = variable1, predhappy = variable2)
  df_list[[i]] <- tmp_df
}

s2_task_dat_long <- do.call(rbind, df_list) %>%
  gather(name, value, -subjID) %>% group_by(subjID, name) %>% mutate(data_point = row_number())

# choose example subject ID = 309
subj_id = 309

s2_task_dat$datatosave[[subj_id]][[13]]
s2_r2 <- format(round(s2_task_dat$datatosave[[subj_id]][[13]], 2), nsmall = 2)

# PHQ-9 score of the example participant 
study2_data$phq[subj_id]

fig_3a <- ggplot(s2_task_dat_long %>% filter(subjID == subj_id), aes(x = data_point, y=value, group=name, linetype=name, shape=name, colour=name)) +
  geom_line(size = 1) + ylim(c(0,100)) +
  geom_point(size = 3) + theme_pubr() + xlab('Rating Number') + ylab('Happiness Rating') +
  scale_linetype_manual(values = c("rawhappy" = "solid", "predhappy" = "solid"), labels = c("rawhappy" = "Raw Happiness", "predhappy" = "Predicted Happiness")) +
  scale_shape_manual( values = c("rawhappy" = 16, "predhappy" = 17), labels = c("rawhappy" = "Raw Happiness", "predhappy" = "Predicted Happiness")) +
  scale_color_manual(values = c("rawhappy" = "blue", "predhappy" = "red"), labels = c("rawhappy" = "Raw Happiness", "predhappy" = "Predicted Happiness")) + theme(legend.text=element_text(size=14), legend.position = c(0.375,0.85), legend.title = element_blank(),
        text = element_text(size = 14), axis.title = element_text(size = 16))

# print Fig. 3A
fig_3a

# draw Fig 3b
# current mood (baseline happiness) by binned PHQ-9
fig_3b <- dat %>%
  group_by(Group = cut(phq, breaks = c(-Inf, 4, 9, 14, 19, +Inf),
                       labels = c("0-4", "5-9", "10-14","15-19", "20+"))) %>%
  summarise(mean_basehap = mean(base_hap), sem_basehap = sd(base_hap) / sqrt(n())) %>%

  # create the bar plot
  ggplot() +
  geom_point(aes(x = Group, y = mean_basehap, fill = Group), stat = "identity", size=1.5,
             position = position_dodge(width = 0.4), width = 1) + geom_hline(yintercept=50, alpha=0.5) +
  geom_errorbar(aes(x = Group, ymin = mean_basehap - sem_basehap, ymax = mean_basehap + sem_basehap), size = 1,
                position = position_dodge(width = 2), width = 0.2) +
  
  # customize plot aesthetics
  labs(x = "Initial PHQ-9 Score",
       y = "Baseline Mood Parameter") +
  theme_pubr() + ylim(c(0,100)) + theme(legend.position="none", text = element_text(size = 14), axis.title = element_text(size = 16))

# print Fig. 3B
fig_3b

# save Fig. 3
fig_3 <- ggarrange(fig_3a, fig_3b, ncol=2, labels = c("A", "B"))
ggsave(filename = file.path(dir_save, "fig_3.pdf"), plot=fig_3, width=9, height=4, units="in", dpi=600)
```

Robust regression testing current depression using human sentiment score 
```{r ms 18, warning = FALSE}
## Study 1
ms18_study1_mod <- lmrob(phq   ~ age + gender + edu + avg_human_tone_mean_diff + base_hap, data=study1_data,
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms18_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms18_study1_mod2<-lmrob(phq   ~  age + gender + edu + avg_human_tone_mean_diff, 
                        data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms18_study1_mod)$r.squared - summary(ms18_study1_mod2)$r.squared)/(1-summary(ms18_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms18_study2_mod <- lmrob(phq   ~ age + gender + edu + avg_human_tone_mean_diff + base_hap,
                         data=study2_data, 
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms18_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms18_study2_mod2<-lmrob(phq   ~  age + gender + edu + avg_human_tone_mean_diff, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms18_study2_mod)$r.squared - summary(ms18_study2_mod2)$r.squared)/(1-summary(ms18_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

Robust regression testing baseline mood parameter  
```{r ms 19, warning = FALSE}
## Study 1
ms19_study1_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap, data=study1_data,
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms19_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without baseline mood par
ms19_study1_mod2<-lmrob(phq_2  ~ phq + age + gender + edu, 
                        data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms19_study1_mod)$r.squared - summary(ms19_study1_mod2)$r.squared)/(1-summary(ms19_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms19_study2_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap,
                         data=study2_data, 
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms19_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without baseline mood par
ms19_study2_mod2<-lmrob(phq_2   ~phq + age + gender + edu, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms19_study2_mod)$r.squared - summary(ms19_study2_mod2)$r.squared)/(1-summary(ms19_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.5f', s2_mod_f)
```

Robust regression testing human sentiment score beyond baseline mood
```{r ms 20, warning = FALSE}
## Study 1
ms20_study1_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap + avg_human_tone_mean_diff, data=study1_data,
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms20_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms20_study1_mod2<-lmrob(phq_2  ~ phq + age + gender + edu + base_hap, 
                        data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms20_study1_mod)$r.squared - summary(ms20_study1_mod2)$r.squared)/(1-summary(ms20_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms20_study2_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap + avg_human_tone_mean_diff,
                         data=study2_data, 
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms20_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms20_study2_mod2<-lmrob(phq_2   ~ phq + age + gender + edu + base_hap, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms20_study2_mod)$r.squared - summary(ms20_study2_mod2)$r.squared)/(1-summary(ms20_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

Robust regression testing GPT sentiment score beyond baseline mood
```{r ms 21, warning = FALSE}
## Study 1
ms21_study1_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap + gpt_tone_mean_diff, data=study1_data,
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms21_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms21_study1_mod2<-lmrob(phq_2  ~ phq + age + gender + edu + base_hap, 
                        data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms21_study1_mod)$r.squared - summary(ms21_study1_mod2)$r.squared)/(1-summary(ms21_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms21_study2_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap + gpt_tone_mean_diff,
                         data=study2_data, 
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms21_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms21_study2_mod2<-lmrob(phq_2   ~ phq + age + gender + edu + base_hap, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms21_study2_mod)$r.squared - summary(ms21_study2_mod2)$r.squared)/(1-summary(ms21_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

Robust regression testing LIWC sentiment score beyond baseline mood
```{r ms 22, warning = FALSE}
## Study 1
ms22_study1_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap + liwc_tone_mean_diff, data=study1_data,
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms22_study1_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms22_study1_mod2<-lmrob(phq_2  ~ phq + age + gender + edu + base_hap, 
                        data=study1_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s1_mod_f <- (summary(ms22_study1_mod)$r.squared - summary(ms22_study1_mod2)$r.squared)/(1-summary(ms22_study1_mod)$r.squared)
sprintf('Study 1 regression model Cohen f^2 = %.3f', s1_mod_f)

## Study 2
ms22_study2_mod <- lmrob(phq_2   ~ phq + age + gender + edu + base_hap + liwc_tone_mean_diff,
                         data=study2_data, 
                         control = lmrob.control(setting = "KS2014", k.max=2000))

tab_model(ms22_study2_mod,  terms = terms_use, 
          show.ci = FALSE, show.se = TRUE,
             pred.labels = pi, dv.labels = c(""), string.pred = "Coeffcient",
            show.stat = TRUE, string.est = "Beta estimates", string.stat = "T-stat",
            string.se = "SE", string.p = "P")

# f^2 calculation
# model without sentiment score
ms22_study2_mod2<-lmrob(phq_2   ~ phq + age + gender + edu + base_hap, data=study2_data,  control = lmrob.control(setting = "KS2014", k.max=2000))
s2_mod_f <- (summary(ms22_study2_mod)$r.squared - summary(ms22_study2_mod2)$r.squared)/(1-summary(ms22_study2_mod)$r.squared)
sprintf('Study 2 regression model Cohen f^2 = %.3f', s2_mod_f)
```

